\Cengine{} is a lightweight compute engine designed to run in the background, below user level \cpp{} code.  
The puropose of \Cengine{} is to dynamically parallelize computations by  
%dynamically parallelizes %unstructured 
%numerical computations in a combination of two ways: 
%mathematical workloads by 
\begin{compactenum}[~~(a)]
\item distributing certain numerical operations across multiple CPU threads and/or  
\item batching together operations of the same kind for parallel execution on the GPU. 
\end{compactenum}
\Cengine{} employs the delayed execution model of computation to decouple user code %calling code 
from the worker threads. 
%give itself freedom in which 
%order to execute the operations demanded by user level code. 
%Internally, it 
It must therefore maintain a internal dependency graph between operations   
to ensure correctness. %that dependency relationships are not violated. 

From the user code side, the engine expects a sequence of simple instructions. For example, 

\texttt{~~~c=engine.push<ctensor\_add\_op>(a,b)}

tells the engine to add tensors \ccode{a} and \ccode{b} and store the result in \ccode{c}. 
Instead of executing this instruction directly, 
%The instructions are not executed immediately. 
%Instead, 
\Cengine{} just adds the corresponding \ccode{ctensor\_add\_op} operator to its internal queue, 
and executes it later, when either a CPU threads becomes 
available or a sufficient number of operations of the same type have accumulated to 
make executing them on the GPU together economical. 
%The result of a sequence of operations is returned to the user once it is ready. 

\Cengine{} offers a small collection of simple built-in data types such as \ccode{rscalar, cscalar} and \ccode{ctensor} 
to store real/complex valued scalar and tensor objects, and a corresponding complement of 
basic arithmetic and linear algebra operators. 
However, the engine is designed to manage user defined data types and 
operators just as well as it can manage its built-in types. 
%, and in fact we expect that in most use cases the engine will be extended in this way. 

\Cengine{} is written in standard \cppe{} and requires no special libraries besides CUDA/CUBLAS 
for GPU functionality.  

